---
title: "Unit Tests"
author: "KF"
output: html_document:
editor_options: 
  chunk_output_type: console
---

## Setup

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r load functions}
source(here("src","functions.R"))
```

### Libraries

For users:

```{r user lib}
gc()

#libraries copied directly from app.R
library(shiny)
library(shinyjs)
library(tidyverse)
library(waiter)
library(DT)
```

For dev:

```{r dev lib}
library(data.table)
library(here)
library(kqol)
```

This code chunk is to avoid having to type out full name of example dataset.

```{r}
getwd()

#store path to example data folder
data_folder <- here("UnitTests","Ex_Data")

#list files inside folder
ex_data_files <- list.files(data_folder)

(ex_data_files)
```

### Sample Data

```{r}
#'imports all files inside specified directory
  #'Package + Documentation: https://github.com/kenf1/KF.QoL
ex_data_list <- kqol::bulkImport(
  FileExtension = ".csv",
  FolderPath = data_folder
)

rm(data_folder,ex_data_files)
```

Alternative import using primarily base R:

```{r}
# ex_data_list <- list()
# for(i in 1:length(ex_data_files)){
#   ex_data_list[[i]] <- read_csv(ex_data_files[i])
# }
# names(ex_data_list) <- ex_data_files
```

## Tests

### validate_odc

This function is a wrapper around the `validate_schema` and `validate_structure` functions.

```{r}
#list to store/organize all unit testing outputs below
outputs <- list()
```

```{r}
outputs[["all"]] <- validate_odc(
  dataset = ex_data_list$`Example ODC dataset`,
  datadic = ex_data_list$`Example ODC data dictionary`,
  str_checks = "all",
  sch_checks = "all"
)
```

entire function as a whole working as expected after removing duplicated `sch_checks` var in `validate_schema` function

#### validate_schema

```{r}
# outputs[["schema"]] <- validate_schema(dataset = ex_data_list$`Example ODC dataset`,
#                  datadic = ex_data_list$`Example ODC data dictionary`)
```

#### validate_structure

#### pos1_char

```{r}
var_names <- ex_data_list$`Example ODC data dictionary`$VariableName

var_names[5] <- "3abc"

#apply check
pos1_check <- lapply(var_names,first_char_letter)
check_temp_df <- data.frame(Key=var_names,Value=unlist(pos1_check))
    
#keep datadic_vars that failed check
check_temp_df <- check_temp_df %>%
      dplyr::filter(Value=="Fail")

rm(var_names,pos1_check,check_temp_df)
```

### checklist_to_df

```{r}
outputs[["ex_df"]] <- checklist_to_df(outputs$all$structure)
head(outputs$ex_df)
```

don't worry about warning message of:
  "In stri_replace_all_regex(string, pattern, fix_replacement(replacement),  :
  argument is not an atomic vector; coercing"

### startApp

`startApp` is only for the package version of this app

```{r}
# startApp(loadReqPackages = FALSE)
```

### run app

to run app (non-package version)

```{r}
source("./app.R")
shinyApp(ui=ui,server=server)
```
